Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-02-06 analysis/paper/paper.Rmd
add cross reference and one figure

We compare the result of shape variable with the result of metric data for six measurements and two ratio values using CV test (Table \@ref(tab:metricCVstest)). Metric measurements show there is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter before and after the European presence. After the arrival of the Chinese, there is only one metric attribute, the rim diameter, showing significant difference compared with post-European period (p-value less than 0.05). Although these results show the differences between different contact phases that corresponds to part of the results of shape variable, elliptic Fourier analysis captures more subtle differences in a substantial way reflected by overall shapes of pottery, not only limited to diameter variables. In addition, results of spatial analysis for pottery shape samples show a non-randomly dispersed distribution before European contact and more extreme dispersed after European presence. However, the distribution of pottery is more similar to the random distributions during the Chinese period. This contradicts our expectation that clustered pattern will be observed with an increase in pottery standardization since the emergence of specialized groups. Or, this indicates that there are other factors lead to pottery standardization instead of a small number of artisans.

>>>>

We compare the result of shape variable with the result of metric data for six measurements and two ratio values using CV test (Table \@ref(tab:metricCVstest)). Metric measurements show there is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter before and after the European presence. After the arrival of the Chinese, there is only one metric attribute, the rim diameter, showing significant difference compared with post-European period (p-value less than 0.05). Although these results show the differences between different contact phases that corresponds to part of the results of shape variable, elliptic Fourier analysis captures more subtle differences in a substantial way reflected by overall shapes of pottery, not only limited to diameter variables. In addition, results of spatial analysis (Figure \@ref(fig:kernel-plot); Figure \@ref(fig:CV-test-PCs-plot)) for pottery shape samples show a non-randomly dispersed distribution before European contact and more extreme dispersed after European presence. However, the distribution of pottery is more similar to the random distributions during the Chinese period. This contradicts our expectation that clustered pattern will be observed with an increase in pottery standardization since the emergence of specialized groups. Or, this indicates that there are other factors lead to pottery standardization instead of a small number of artisans.

```{r kernel-plot, fig.cap= "Kernel density map visulizes the probability of the density of pottery across space. The map shows  major core area during the pre-European period, multiple core areas during the European period, and a single core during the Chinese period. Used the bandwidth based on Silverman (1986)'s rule of thumb"}
knitr::include_graphics(here("analysis", "figures", "plot-kde-maps.jpg"))
```
